---
title: "Convert ArchR to Signac"
author: "kou"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, include=T, message=F)
```

## Converting ArchR to Signac Chromatin Assay
Note: some "items" can be calculated within Signac, such as Tn5 insertion bias matrix. Can always add additional matrices calculated in ArchR.
```{r libraries}
library(ArchRtoSignac)
library(ArchR)
library(Signac)
library(Seurat)
library(stringr)
library(EnsDb.Hsapiens.v86)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(purrr)
library(qs)

set.seed(105)
addArchRThreads(threads = 32, force = T)
options(future.globals.maxSize = 4000 * 1024^2)
```

Load ArchR projects
```{r load projects}
dir = paste0("01_QC/ArchR_",c("LN0025","LN0177","LN0193","LN0438"))

archR_all = lapply(dir,loadArchRProject)
```

Get path to fragment files
```{r fragments}
# list of samples
ln_id = c("LN0025","LN0177","LN0193","LN0438")

dir = paste0("~/projects/composite/cellranger/",ln_id,"/outs/")

# get vector of frag files
fragmentFiles = list.files(path=dir, full.names = T, pattern = "fragments.tsv.gz$")
```

### Get counts of binned genome
Need to use Tile Matrix that has NOT binarized. 
Had to repeat ATAC seq processing to generate new arrow Files and ArchR project with non-binarized matrix. 
Note on ranges: https://github.com/GreenleafLab/ArchR/issues/330
```{r test1, eval = F}
# practice to understand how to get info
## get SE
test = getMatrixFromProject(archR_all[[1]],"TileMatrix")

## get unbin sparse matrix w/ rows = insertion frags, cols = cells
assay(test)

## get ranges
rowData(test)

## get cell metadata
colData(test)

## get cell names
head(colnames(test))
```

```{r get count matrix}
# get list of SE containing tile matrix
# TO DO: rename columns here to shorten code! 
SE_nonBin = lapply(archR_all, getMatrixFromProject, useMatrix = "TileMatrix")

counts_nonBin = lapply(SE_nonBin, assay)
names(counts_nonBin) = ln_id

# rename colnames so matches original fragment file (ie: only cell barcode, no "LNXXXX#" prefix)
counts_nonBin = lapply(counts_nonBin, function(x){
  colnames(x) = gsub("LN.*#","",colnames(x)) 
  return(x)
})
```

### Get ranges assocated with binned genome
```{r test2, eval=F}
# test granges conversion
testGR = rangesOfCounts[[1]]
testGR$end = testGR$start + 500 - 1
testGR = makeGRangesFromDataFrame(testGR)

# check none are overlapping
isDisjoint(testGR)

# check bin size all correct
all(width(testGR)==500)
```

```{r get ranges}
# get ranges, convert to GRanges
rangesOfCounts = lapply(SE_nonBin, rowData)

# create function to make granges
convert2GR = function(rowDataDF){
  
  # end is binsize - 1 (ie: 500 - 1)
  rowDataDF$end = rowDataDF$start + 499 
  rowDataGR = makeGRangesFromDataFrame(rowDataDF)
  return(rowDataGR)
}

# convert ranges
rangesOfCounts = lapply(rangesOfCounts,convert2GR)

# since using TileMatrix to create chromatinAssay, the binned genome is the same for all samples
# check w/ lapply(counts_nonBin,dim)
rangesBinned = rangesOfCounts[[1]]
```


### Get Annotation
TO DO: check default archR annotation, perhaps it is w/ TxDb
```{r annotation}
# "UCSC" is the default style to change to but can be changed with argument seqStyle
# confirmed that this is the correct annotation style for signac based on signac vignette
# used wrapper function from ArchRtoSignac
hg38_anno = getAnnotation(reference = EnsDb.Hsapiens.v86, refversion = "hg38", seqStyle = "UCSC")
```

### Create Chromatin Assays
Notes on structure: https://stuartlab.org/signac/articles/data_structures.html
```{r create chromatin assays}
# number of samples may change, so use seq_along
chrAssays = map(seq_along(counts_nonBin), ~ CreateChromatinAssay(
  counts = counts_nonBin[[.x]],
  fragments = fragmentFiles[[.x]],
  ranges = rangesBinned,
  genome = "hg38",
  annotation = hg38_anno,
  validate.fragments = T
))

names(chrAssays) = ln_id
```

### Convert to Seurat Object for integration
Create `Seurat Object` for integration in next step.
Note: Had to convert now to rename cells (did not work for `Chromatin Assay` object at the time of processing).
```{r seur obj}
# first get metadata to add to seurat object
metadata_list = lapply(archR_all, function(x) x@cellColData)

# rename rows to match names of cells in chrAssay
metadata_list = lapply(metadata_list, function(x){
  newNames = gsub("LN.*#","",rownames(x))
  rownames(x) = newNames
  return(as.data.frame(x))
})

# confirm cells are in same order as in chrAssay
lapply(seq_along(metadata_list), function(x){
  isSameOrder = rownames(metadata_list[[x]]) == Cells(chrAssays[[x]])
  all(isSameOrder)
})

# create Seurat objects
atacSeurat = map2(chrAssays, metadata_list, ~ CreateSeuratObject(
  counts = .x,
  meta.data = .y,
  assay = "ATAC",
  project = "composite_LN"
))

# add prefix to cells
atacSeurat = map2(atacSeurat,ln_id, ~ RenameCells(.x, add.cell.id = .y))
```

### Add Embeddings from ArchR Projects
- UMAP
- Iterative LSI4
```{r add umap}
# get umap embedding
umapEmbed = lapply(archR_all, getEmbedding, embedding = "atacUMAP")

# rename colnames and rownames
umapEmbed = lapply(umapEmbed, function(x){
  colnames(x) = c("atacUMAP_1","atacUMAP_2")
  rownames(x) = gsub("#","_",rownames(x))
  return(as.matrix(x))
})

# get LSI embedding
lsi4Dim = lapply(archR_all, getReducedDims, reducedDims = "LSI_4")

# rename rownames
lsi4Dim = lapply(lsi4Dim, function(x){
  rownames(x) = gsub("#","_",rownames(x))
  return(as.matrix(x))
})

# add embeddings to seurat object
atacSeurat = lapply(seq_along(atacSeurat), function(x){
  atacSeurat[[x]][["atacUMAP"]] = CreateDimReducObject(embeddings = umapEmbed[[x]], key = "atacUMAP_", assay = "ATAC")
  atacSeurat[[x]][["iterLSI_4"]] = CreateDimReducObject(embeddings = lsi4Dim[[x]], key = "atacIterLSI4_", assay = "ATAC")
  return(atacSeurat[[x]])
})
```

### Add Gene Score Activity Matrix
```{r get gene score matrix}
# get imputed gene score to add to chromatin assay
geneScoreMat = lapply(archR_all, getMatrixFromProject, useMatrix = "GeneScoreMatrix", verbose=F)
geneNames = geneScoreMat[[1]]@elementMetadata$name

# get impute weights
imputeWeights = lapply(archR_all, getImputeWeights)

# get matrix w impute scores, need to use se object in geneScoreMat to get gene names (rows)
imputedMat = map2(geneScoreMat,imputeWeights, ~ imputeMatrix(assay(.x),.y))
names(imputedMat) = ln_id

# rename rows
imputedMat = lapply(imputedMat, function(x){
  rownames(x) = geneNames
  colnames(x) = gsub("#","_",colnames(x)) 
  
  # round
  x = round(x, digits = 3)
  
  # convert to matrix
  x = as.matrix(x)
  return(x)
})

# add imputed gene score activity as an assay for access later
## cols = cells, rows = features
atacSeurat = lapply(seq_along(atacSeurat), function(x){
  atacSeurat[[x]][["GeneScoreImputed"]] = CreateAssayObject(data = imputedMat[[x]])
  return(atacSeurat[[x]])
})
```

save object for integration
```{r save qs}
qsave(atacSeurat,"02_integration/rds/01_atac_seurat.qs",nthreads = 16)
#atacSeurat = qread("02_integration/rds/01_atac_seurat.qs")
```