---
title: "sct integration"
author: "kou"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,include=T,message=F,eval=F,warning=F)
```

```{r libraries, warning=F}
library(Signac)
library(Seurat)
library(tidyverse)
library(ggpubr)
library(shiny)
library(qs)
library(reticulate)
library(clustree)

# always set seed!!! need to repeat
set.seed(105)

options(future.globals.maxSize = 4000 * 1024^2)

# set python
use_python("/g/scb/zaugg/kou/projects/composite/analysis/sc_py_env/bin")
use_virtualenv("/g/scb/zaugg/kou/projects/composite/analysis/sc_py_env")
``` 

```{r data_import}
ln_id = c("LN0025","LN0177","LN0193","LN0438")
sct_list = qread("01_QC/rds/04_split_sct_noCCreg.qs")
```

## Integrate through finding anchors and batch correct SCT modalities
Samples were first normalized independently using SC Transform v2 (see 01_QC/04_gex_norm_final.Rmd), then integrated using pearson residuals.
3000 features were used as anchors for integration. Including more anchors is not necessarily beneficial, as the additional feautures would have lower weights.

No reference was assigned, so all pairwise anchors are found between all samples.

More info: https://satijalab.org/seurat/articles/sctransform_v2_vignette.html
```{r integrate_anchors}
# reformat metadata before integration
sct_list = lapply(1:4, function(x){
  # get metadata from sct_list
  meta = sct_list[[x]]@meta.data
  
  # get vector with resolutions
  res = str_extract(colnames(meta),"res.*")
  
  # get col index with RNA_snn_res
  index_rna_res = str_which(colnames(meta),"RNA_snn_res.")
  
  # get col index with SCT_snn_res 
  index_sct_res = str_which(colnames(meta),"SCT_snn_res.")
  
  # rename columns
  meta = meta %>%
    mutate(RNA_snn_res = res[index_rna_res],
           SCT_snn_res = res[index_sct_res]) %>%
    rename(RNA_snn_clust = colnames(meta)[index_rna_res],
           SCT_snn_clust = colnames(meta)[index_sct_res])
  
  sct_list[[x]]@meta.data = meta
  
return(sct_list[[x]])
})

# ranks features by the number of datasetes that they're variable in, returns top scoring features
features = SelectIntegrationFeatures(object.list = sct_list, 
                                     nfeatures = 3000, 
                                     assay = rep("SCT",length(sct_list)))

sct_list = PrepSCTIntegration(object.list = sct_list,
                              assay = "SCT", 
                              anchor.features = features)

# integrate the datasets
anchors = FindIntegrationAnchors(object.list = sct_list, 
                                 assay = rep("SCT",length(sct_list)), 
                                 normalization.method = "SCT", 
                                 anchor.features = features)

sct_integrated = IntegrateData(anchorset = anchors, 
                               normalization.method = "SCT")
```

### Dimensional Reduction
After integrating data sets, a new "integrated" assay was created, and all reductions were removed. Original RNA and SCT counts are also stored. Metadata seems to be rbind() of metadata from list of Seurat objects. 

Next, perform dimensionality reduction to cluster cells and identify marker genes on the "integrated" assay.

Note: set python interpreter to venv through project options, also used reticulate in set up.
```{r dim_reduc, warning=F}
# create umap on sct
sct_integrated = RunPCA(sct_integrated, 
                        reduction.name = "sctPCA", 
                        verbose=F) %>%
  RunUMAP(., reduction = "sctPCA", 
          dims = 1:30, 
          reduction.key = "sctUMAP_", 
          reduction.name = "sctUMAP", 
          verbose=F) %>% 
  FindNeighbors(., reduction = "sctPCA", 
                dims = 1:30)

# use clustree to find the appropriate resolution, cluster with leiden
clustree = FindClusters(sct_integrated, 
                        resolution = seq(0.4,1.15,by = 0.25), 
                        algorithm = 4, 
                        method = "igraph", 
                        random.seed = 1)

# plot to find appropriate resolution
clustree(clustree@meta.data, prefix = "integrated_snn_res.")


# clustering with 0.4
sct_integrated = FindClusters(sct_integrated, 
                              resolution = 0.4, 
                              algorithm = 4, 
                              method = "igraph", 
                              random.seed = 1)
```

```{r save}
qsave(sct_integrated, "02_integration/rds/02_sct_integrated_noCCreg.qs", nthreads = 16)
sct_integrated = qread("02_integration/rds/02_sct_integrated_noCCreg.qs")
```

### Visualizing embeddings
Subset RNA features to color code embeddings
```{r features}
# genes of interest
genes = c("MS4A1",
          "PAX5",
          "CXCR5",
          "CD3D",
          "CD4",
          "CD8A",
          "FOXP3",
          "TBX21",
          "GATA3",
          "BCL6",
          "PDCD1",
          "HAVCR2",
          "BCL2",
          "MKI67")

# get norm counts
normcounts = sct_integrated@assays$RNA@data
normcounts = normcounts[genes,]
```

UMAP: All samples in same embedding
```{r umap}
sctUMAPembed = as.data.frame(Embeddings(sct_integrated,"sctUMAP"))

sctUMAPembed$sample = sct_integrated$Sample

# plot
sctUMAP_allSamples = ggplot(sctUMAPembed, aes(sctUMAP_1,sctUMAP_2)) +
  geom_jitter(aes(color = sample), size = 0.15, show.legend = F) +
  theme_bw() +
  theme(aspect.ratio = 3/3,
        panel.grid = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank())

ggsave("02_integration/figs/03_sctUMAP_allSamples.png", sctUMAP_allSamples, width = 5, height = 5)
```

UMAP: Clusters (integrated_snn_res.0.4) faceted by samples. 
```{r umapClusters}
sctUMAPembed$clusters = sct_integrated$integrated_snn_res.0.4

# plot
sctUMAP_clusters = ggplot(sctUMAPembed, aes(sctUMAP_1,sctUMAP_2)) +
  geom_jitter(aes(color = clusters), size = 0.05, show.legend = T) +
  theme_bw() +
  theme(aspect.ratio = 3/3,
        panel.grid = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank()) +
  facet_grid(cols = vars(sample))

ggsave("02_integration/figs/03_sctUMAP_clusters.png", sctUMAP_clusters, width = 10, height = 5)
```

UMAP: Facet by samples, color by normalized RNA counts.
```{r umapFeature}
# join, after confirming all(rownames(sctUMAPembed) == colnames(normcounts))
sctUMAPembed = cbind(sctUMAPembed, t(as.matrix(normcounts)))

# create a function for umap plotting
umapFeature = function(df_embed,locus){
  ggplot(df_embed, aes(sctUMAP_1,sctUMAP_2)) +
  geom_jitter(aes(color = .data[[locus]]), size = 0.05) +
  scale_color_viridis_c(option = "turbo") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        aspect.ratio = 3/3,
        strip.background = element_blank()) +
  facet_grid(cols = vars(sample))
}

# save, TO DO: lapply
ms4a1 = umapFeature(sctUMAPembed, "MS4A1")
cd3d = umapFeature(sctUMAPembed, "CD3D")
bcl6 = umapFeature(sctUMAPembed, "BCL6")
bcl2 = umapFeature(sctUMAPembed, "BCL2")

ggsave("02_integration/figs/03_sctUMAP_MS4A1.png", ms4a1, width = 10, height = 5)
ggsave("02_integration/figs/03_sctUMAP_CD3D.png", cd3d, width = 10, height = 5)
ggsave("02_integration/figs/03_sctUMAP_BCL6.png", bcl6, width = 10, height = 5)
ggsave("02_integration/figs/03_sctUMAP_BCL2.png", bcl2, width = 10, height = 5)
```

PCA: clusters faceted by samples. 
```{r pca}
sctPCAembed = as.data.frame(Embeddings(sct_integrated,"sctPCA"))

sctPCAembed$sample = sct_integrated$Sample

sctPCAembed$clusters = sct_integrated$integrated_snn_res.0.4

# pc 1 vs pc 2
sctPCA_clusters = ggplot(sctPCAembed, aes(PC_1,PC_2)) +
  geom_jitter(aes(color = clusters), size = 0.05, show.legend = T) +
  theme_bw() +
  theme(aspect.ratio = 3/3,
        panel.grid = element_blank()) +
  facet_grid(cols = vars(sample))

# pc 3 vs pc 4
sctPCA_clusters2 = ggplot(sctPCAembed, aes(PC_3,PC_4)) +
  geom_jitter(aes(color = clusters), size = 0.05, show.legend = T) +
  theme_bw() +
  theme(aspect.ratio = 3/3,
        panel.grid = element_blank()) +
  facet_grid(cols = vars(sample))

ggsave("02_integration/figs/03_sctPCA_PC1PC2_clusters.png", sctPCA_clusters, width = 10, height = 5)
ggsave("02_integration/figs/03_sctPCA_PC3PC4_clusters.png", sctPCA_clusters2, width = 10, height = 5)
```

PCA: Facet by samples, color by normalized RNA counts (same matrix used for UMAP). 
```{r pcaFeature}
# join, after confirming all(rownames(sctUMAPembed) == colnames(normcounts))
sctPCAembed = cbind(sctPCAembed, t(as.matrix(normcounts)))

# create a function for pca plotting
pcaFeature = function(df_embed,locus,whichPC1,whichPC2){
  ggplot(df_embed, aes(.data[[whichPC1]],.data[[whichPC2]])) +
  geom_jitter(aes(color = .data[[locus]]), size = 0.05) +
  scale_color_viridis_c(option = "turbo") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        aspect.ratio = 3/3,
        strip.background = element_blank()) +
  facet_grid(cols = vars(sample))
}

# save, TO DO: lapply or shiny app
ms4a1_pca = pcaFeature(sctPCAembed, "MS4A1","PC_1","PC_2")
cd3d_pca = pcaFeature(sctPCAembed, "CD3D","PC_1","PC_2")
mki67_pca = pcaFeature(sctPCAembed, "MKI67","PC_3","PC_4")

ggsave("02_integration/figs/03_sctPCA_MS4A1.png", ms4a1_pca, width = 10, height = 5)
ggsave("02_integration/figs/03_sctPCA_CD3D.png", cd3d_pca, width = 10, height = 5)
ggsave("02_integration/figs/03_sctPCA_MKI67.png", mki67_pca, width = 10, height = 5)
```